---
title: "IE582 Fall"
author: "Fatma Nur Dumlupınar"
date: "29 01 2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    code_folding: hide
    theme: journal
  pdf_document:
    toc: yes
    toc_depth: '3'
subtitle: Homework 4
---

```{r message=FALSE, warning=FALSE}
library(data.table)
library(farff)
library(glmnet)
library(caTools)
library(caret)
library(e1071)
library(MLmetrics)
library(dplyr)

```


# Online News Popularity Data Set

```{r message=FALSE, warning=FALSE}
OnlineNews<-read.csv("data/Online News Popularity Data Set/OnlineNewsPopularity.csv") #categorical,regression,61
OnlineNews<-OnlineNews[38644:39644,]

OnlineNews<-OnlineNews[,-1]
OnlineNews<-OnlineNews[,-1]
OnlineNews$shares<-(scale(OnlineNews$shares))
spl=sample.split(OnlineNews$shares, SplitRatio = 0.8)
OnlineNewsTrain=subset(OnlineNews,spl==TRUE)
OnlineNewsTest=subset(OnlineNews,spl==FALSE)
str(OnlineNews)

```

## Penalized Regression Approaches (RPA)


```{r message=FALSE, warning=FALSE}
###PRA###########################################################

PRAmodel_ONews<- cv.glmnet(as.matrix(OnlineNewsTrain[,1:58]),
                           OnlineNewsTrain$shares,
                           alpha = 1,nfolds=10,
                           family="gaussian")
PRAmodel_ONews

plot(PRAmodel_ONews)

ONewsPredicted_RPA_Train<-predict(PRAmodel_ONews,as.matrix(OnlineNewsTrain[,1:58]),
                            s=PRAmodel_ONews$lambda.min)

ONewsPredicted_RPA<-predict(PRAmodel_ONews,as.matrix(OnlineNewsTest[,1:58]),
                            s=PRAmodel_ONews$lambda.min)

CVError_PRA_ONews<-min(PRAmodel_ONews$cvm)

RMSE_PRA_ONews<-RMSE(OnlineNewsTest$shares,ONewsPredicted_RPA)
RMSE_PRA_ONews_Train<-RMSE(OnlineNewsTrain$shares,ONewsPredicted_RPA_Train)

```

## Random Forests (RF)

```{r message=FALSE, warning=FALSE}
###RF#########################################################


fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid")    

RFtuneGrid=expand.grid(mtry = c(5,15,25))
# the minimal number of observations per tree leaf=5, number of tree J=500,
RFmodel_ONews <- train( shares~ ., 
                        data=as.matrix(OnlineNewsTrain), 
                        method="rf",
                        metric="RMSE",
                        ntree =500,nodesize=5,
                        trControl=fitControl,
                        tuneGrid = RFtuneGrid)

ONewsPredicted_RF<-predict(RFmodel_ONews,as.matrix(OnlineNewsTest))

CVError_RF_ONews<-min(RFmodel_ONews$results$RMSE)

RMSE_RF_ONews<-RMSE(OnlineNewsTest$shares,ONewsPredicted_RF)

print(RFmodel_ONews)


```

## Decision Trees (DT)

```{r message=FALSE, warning=FALSE}
#############Decision Tree###############

#Online News

fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid") 

DTtuneGrid<-expand.grid(cp = c(0.001,0.01, 0.1))

Results_DT_ONews<-data.frame()

minbucket=0

for (minbucket in c(5,10,15)){
  
  DTmodel_ONews <- train(shares ~ ., 
                           data=as.matrix(OnlineNewsTrain), 
                           method="rpart", 
                           metric="RMSE",
                           minbucket =minbucket, 
                           trControl=fitControl,
                           tuneGrid = DTtuneGrid)
  
  RMSECpMinbucket<-DTmodel_ONews$results[,c("RMSE","cp")]
  RMSECpMinbucket$minbucket<-minbucket
  Results_DT_ONews<-rbind(RMSECpMinbucket,Results_DT_ONews)
  
  
}

res<-Results_DT_ONews%>%arrange((RMSE))%>%head(1)
res

DTtuneGrid <-  expand.grid(cp =res$cp)

DTmodel_ONews_best <- train(shares ~ ., 
                       data=as.matrix(OnlineNewsTrain), 
                       method="rpart", 
                       metric="RMSE",
                       minbucket =res$minbucket, 
                       trControl=fitControl,
                       tuneGrid = DTtuneGrid)



ONewsPredicted_DT<-predict(DTmodel_ONews_best,as.matrix(OnlineNewsTest))

CVError_DT_ONews<-min(DTmodel_ONews_best$results$RMSE)

RMSE_DT_ONews<-RMSE(OnlineNewsTest$shares,ONewsPredicted_DT)
```

## Stochastic Gradient Boosting (SGB)


```{r message=FALSE, warning=FALSE}
#############################GBM###################################

fitcontrol =trainControl(method="cv", number=5, search="grid")

GBMtuneGrid=expand.grid(interaction.depth = c(2, 4, 6), 
                        shrinkage = c(0.05,0.1, 0.15),
                        n.trees = c(100, 150, 250),
                        n.minobsinnode = 15)

GBMmodel_ONews <- train(shares ~ ., 
                        data=as.matrix(OnlineNewsTrain), 
                        method="gbm", 
                        metric="RMSE", 
                        trControl=fitcontrol,
                        tuneGrid = GBMtuneGrid)


ONewsPredicted_GBM<-predict(GBMmodel_ONews,OnlineNewsTest)


TRMSE_GBM_ONews<-min(GBMmodel_ONews$results$RMSE)
RMSE_GBM_ONews<-RMSE(ONewsPredicted_GBM,OnlineNewsTest$shares)

GBMmodel_ONews$bestTune
#print(as.matrix(GBMmodel_HDigits$results))
plot(GBMmodel_ONews)
```


## Comparison of the Results

**PRA**

*Cross Validation Error*

```{r message=FALSE, warning=FALSE}
CVError_PRA_ONews

```

*RMSE-Train*


```{r message=FALSE, warning=FALSE}
RMSE_PRA_ONews_Train
```

*RMSE-Test*

```{r message=FALSE, warning=FALSE}
RMSE_PRA_ONews
```


*Lambda*

```{r}
PRAmodel_ONews$lambda.min
```

**RF**

*RMSE-Train* 

```{r message=FALSE, warning=FALSE}
RMSE_PRA_ONews_Train
```

*Cross Validation Error*

```{r message=FALSE, warning=FALSE}
CVError_RF_ONews
```

*RMSE-Test* 

```{r message=FALSE, warning=FALSE}
RMSE_RF_ONews

```

*m*

```{r}
RFmodel_ONews$bestTune
```

**DT**

*RMSE-Train* 

```{r message=FALSE, warning=FALSE}
CVError_DT_ONews

```

*RMSE-Test* 


```{r message=FALSE, warning=FALSE}
RMSE_DT_ONews
```

*Cp and Minbucket*

```{r}
res$cp
res$minbucket
```

**SGB**
 
*RMSE-Train* 

```{r message=FALSE, warning=FALSE}
TRMSE_GBM_ONews
```

*RMSE-Test* 

```{r message=FALSE, warning=FALSE}
RMSE_GBM_ONews
```

*Depth, Learning Rate and Number of Trees*

```{r}
GBMmodel_ONews$bestTune
```

