---
title: "IE582 Fall"
author: "Fatma Nur Dumlupınar"
date: "29 01 2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    code_folding: hide
    theme: journal
  pdf_document:
    toc: yes
    toc_depth: '3'
subtitle: Homework 4
---



```{r message=FALSE, warning=FALSE}
library(data.table)
library(farff)
library(glmnet)
library(caTools)
library(caret)
library(e1071)
library(MLmetrics)
library(dplyr)

```


# Breast Cancer Wisconsin (Diagnostic) Data Set

```{r message=FALSE, warning=FALSE}
BreastCancer<-fread("data/Breast Cancer Wisconsin (Diagnostic)/wdbc.txt")
BreastCancer$V1<-as.numeric(BreastCancer$V1)
BreastCancer$V2<-as.factor(BreastCancer$V2)


set.seed(150)
spl=sample.split(BreastCancer$V2, SplitRatio = 0.8)
BreastCancerTrain=subset(BreastCancer,spl==TRUE)
BreastCancerTest=subset(BreastCancer,spl==FALSE)
BreastCancerTrain<-BreastCancerTrain[,-1]
BreastCancerTest<-BreastCancerTest[,-1]

BreastCancerTrainTarget<-BreastCancerTrain$V2
BreastCancerTestTarget<-BreastCancerTest$V2
BreastCancerTrain<-scale(BreastCancerTrain[,-1])
BreastCancerTest<-scale(BreastCancerTest[,-1])
BreastCancerTest<-cbind(BreastCancerTestTarget,BreastCancerTest)
BreastCancerTrain<-cbind(BreastCancerTrainTarget,BreastCancerTrain)
BreastCancerTest<-as.data.frame(BreastCancerTest)
BreastCancerTrain<-as.data.frame(BreastCancerTrain)

table(factor(BreastCancer$V2))

str(BreastCancerTrain)

```

## Penalized Regression Approaches (RPA)

```{r message=FALSE, warning=FALSE}
PRAmodel_BreastC<- cv.glmnet(as.matrix(BreastCancerTrain[,-1]),
                             as.factor(BreastCancerTrain$BreastCancerTrainTarget), 
                             alpha = 1,nfolds=10,
                             type.measure='class',family="binomial")
PRAmodel_BreastC

plot(PRAmodel_BreastC)


BreastCPredicted_RPA_Train<-predict(PRAmodel_BreastC,as.matrix(BreastCancerTrain[,-1]),
                              type = "class",s=PRAmodel_BreastC$lambda.min)
BreastCPredicted_RPA<-predict(PRAmodel_BreastC,as.matrix(BreastCancerTest[,-1]),
                              type = "class",s=PRAmodel_BreastC$lambda.min)

CVError_PRA_BreastC<-min(PRAmodel_BreastC$cvm)

CM_PRA_BreastC<-table(as.factor(BreastCancerTest$BreastCancerTestTarget),BreastCPredicted_RPA)

Acc_PRA_BreastC <- sum(diag(CM_PRA_BreastC)) / sum(CM_PRA_BreastC)

CM_PRA_BreastC_Train<-table(as.factor(BreastCancerTrain$BreastCancerTrainTarget),BreastCPredicted_RPA_Train)

Acc_PRA_BreastC_Train <- sum(diag(CM_PRA_BreastC_Train)) / sum(CM_PRA_BreastC_Train)


```

## Random Forests (RF)


```{r message=FALSE, warning=FALSE}
fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid")    

RFtuneGrid=expand.grid(mtry = c(5,15,25))

RFmodel_BreastC <- train(as.factor(BreastCancerTrainTarget) ~ ., 
                         data=as.matrix(BreastCancerTrain), 
                         method="rf",
                         metric="Accuracy",
                         ntree =500,nodesize=5,
                         trControl=fitControl,
                         tuneGrid = RFtuneGrid)

BreastCPredicted_RF<-predict(RFmodel_BreastC,as.matrix(BreastCancerTest))

TAcc_RF_BreastC<-max(RFmodel_BreastC$results$Accuracy)

CM_RF_BreastC<-table(as.factor(BreastCancerTest$BreastCancerTestTarget),BreastCPredicted_RF)
Acc_RF_BreastC <- sum(diag(CM_RF_BreastC)) / sum(CM_RF_BreastC)
```

## Decision Trees (DT)

```{r message=FALSE, warning=FALSE}
###################DT##############################



fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid") 

minbucket=0
DTtuneGrid<-expand.grid(cp = c(0.001,0.01, 0.1))

Results_DT_BreastC<-data.frame()

for (minbucket in c(5,10,15)){
  
  DTmodel_BreastC <- train(factor(BreastCancerTrainTarget) ~ ., 
                           data=as.matrix(BreastCancerTrain), 
                           method="rpart", 
                           metric="Accuracy",
                           minbucket =minbucket, 
                           trControl=fitControl,
                           tuneGrid = DTtuneGrid)
  
  AccCpMinbucket<-DTmodel_BreastC$results[,c("Accuracy","cp")]
  AccCpMinbucket$minbucket<-minbucket
  Results_DT_BreastC<-rbind(AccCpMinbucket,Results_DT_BreastC)
}

res<-Results_DT_BreastC%>%arrange(desc(Accuracy))%>%head(1)
res

DTtuneGrid <-  expand.grid(cp =res$cp)

DTmodel_BreastC_best <- train(factor(BreastCancerTrainTarget) ~ ., 
                           data=as.matrix(BreastCancerTrain), 
                           method="rpart", 
                           metric="Accuracy",
                           minbucket =res$minbucket, 
                           trControl=fitControl,
                           tuneGrid = DTtuneGrid)



BreastCPredicted_DT<-predict(DTmodel_BreastC_best,as.matrix(BreastCancerTest))

TAcc_DT_BreastC<-min(DTmodel_BreastC_best$results$Accuracy)

CM_DT_BreastC<-table(as.factor(BreastCancerTest$BreastCancerTestTarget),BreastCPredicted_DT)

Acc_DT_BreastC<- sum(diag(CM_DT_BreastC)) / sum(CM_DT_BreastC)


```

## Stochastic Gradient Boosting (SGB)


```{r message=FALSE, warning=FALSE}
###################GBM##############################

fitcontrol =trainControl(method="cv", number=5, search="grid")

GBMtuneGrid=expand.grid(interaction.depth = c(2, 4, 6), 
                        shrinkage = c(0.05,0.1, 0.15),
                        n.trees = c(100, 150, 250),
                        n.minobsinnode = 15)

GBMmodel_BreastC <- train(as.factor(BreastCancerTrainTarget) ~ ., 
                          data=as.matrix(BreastCancerTrain), 
                          method="gbm", 
                          metric="Accuracy", 
                          trControl=fitcontrol,
                          tuneGrid = GBMtuneGrid)

BreastCPredicted_GBM<-predict(GBMmodel_BreastC,BreastCancerTest)

TAcc_GBM_BreastC<-max(GBMmodel_BreastC$results$Accuracy)

CM_GBM_BreastC<-table(as.factor(BreastCancerTest$BreastCancerTestTarget),BreastCPredicted_GBM)
Acc_GBM_BreastC <- sum(diag(CM_GBM_BreastC)) / sum(CM_GBM_BreastC)
GBMmodel_BreastC$bestTune
#print(as.matrix(GBMmodel_HDigits$results))
plot(GBMmodel_BreastC)


```


## Comparison of the Results

**PRA**

*Cross Validation Error*

```{r message=FALSE, warning=FALSE}
CVError_PRA_BreastC

```

*Accuracy-Train*

```{r message=FALSE, warning=FALSE}
Acc_PRA_BreastC_Train
```


*Accuracy-Test*

```{r message=FALSE, warning=FALSE}
Acc_PRA_BreastC
```

*Lambda*

```{r}
PRAmodel_BreastC$lambda.min
```

**RF**

*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_RF_BreastC
```

*Accuracy-Test* 

```{r message=FALSE, warning=FALSE}
Acc_RF_BreastC

```

*m*

```{r}
RFmodel_BreastC$bestTune
```

**DT**

*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_DT_BreastC

```

*Accuracy-Test* 


```{r message=FALSE, warning=FALSE}
Acc_DT_BreastC
```

*Cp and Minbucket*

```{r}
res$cp
res$minbucket
```

**SGB**
 
*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_GBM_BreastC
```

*Accuracy-Test* 

```{r message=FALSE, warning=FALSE}
Acc_GBM_BreastC
```

*Depth, Learning Rate and Number of Trees*

```{r}
GBMmodel_BreastC$bestTune
```

