---
title: "IE582 Fall"
author: "Fatma Nur Dumlupınar"
date: "29 01 2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    code_folding: hide
    theme: journal
  pdf_document:
    toc: yes
    toc_depth: '3'
subtitle: Homework 4
---

```{r message=FALSE, warning=FALSE}
library(data.table)
library(farff)
library(glmnet)
library(caTools)
library(caret)
library(e1071)
library(MLmetrics)
library(dplyr)

```


# Cervical Cancer Data Set



```{r message=FALSE, warning=FALSE}
CervicalCancer<-read.csv("data/CervicalCancer/risk_factors_cervical_cancer.csv")  #imbalance
CervicalCancer<-CervicalCancer[,-c(33,34,35)]

for(i in 1:32){
  CervicalCancer[,i]<-as.numeric(CervicalCancer[,i])
  CervicalCancer[is.na(CervicalCancer[,i]), i] <- mean(CervicalCancer[,1], na.rm = TRUE)
}

str(CervicalCancer)

#####Cervical - imbalanced

spl=sample.split(CervicalCancer$Biopsy, SplitRatio = 0.75)
CervicalCancerTrain=subset(CervicalCancer,spl==TRUE)
CervicalCancerTest=subset(CervicalCancer,spl==FALSE)
table(factor(CervicalCancer$Biopsy))

```

## Penalized Regression Approaches (RPA)

```{r message=FALSE, warning=FALSE}
####PRA######################################

PRAmodel_CervicalC<- cv.glmnet(as.matrix(CervicalCancerTrain[,-33]),as.factor(CervicalCancerTrain$Biopsy), 
                               alpha = 1,nfolds=10,
                               type.measure='class',family="binomial")
PRAmodel_CervicalC

plot(PRAmodel_CervicalC)

CervicalCPredicted_RPA<-predict(PRAmodel_CervicalC,as.matrix(CervicalCancerTest[,-33]),
                                type = "class",s=PRAmodel_CervicalC$lambda.min)

CervicalCPredicted_RPA_Train<-predict(PRAmodel_CervicalC,as.matrix(CervicalCancerTrain[,-33]),
                                type = "class",s=PRAmodel_CervicalC$lambda.min)


CVError_PRA_CervicalC<-min(PRAmodel_CervicalC$cvm)

CM_PRA_CervicalC<-table(as.factor(CervicalCancerTest$Biopsy),CervicalCPredicted_RPA)

Acc_PRA_CervicalC <- sum(diag(CM_PRA_CervicalC)) / sum(CM_PRA_CervicalC)

CM_PRA_CervicalC_Train<-table(as.factor(CervicalCancerTrain$Biopsy),CervicalCPredicted_RPA_Train)

Acc_PRA_CervicalC_Train <- sum(diag(CM_PRA_CervicalC_Train)) / sum(CM_PRA_CervicalC_Train)

```

## Random Forests (RF)

```{r message=FALSE, warning=FALSE}
######################RF#####################


fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid")    

RFtuneGrid=expand.grid(mtry = c(5,15,25))
# the minimal number of observations per tree leaf=5, number of tree J=500,

RFmodel_CervicalC <- train(as.factor(Biopsy) ~ ., 
                           data=as.matrix(CervicalCancerTrain), 
                           method="rf",
                           metric="Accuracy",
                           ntree =500,nodesize=5,
                           trControl=fitControl,
                           tuneGrid = RFtuneGrid)

CervicalCPredicted_RF<-predict(RFmodel_CervicalC,as.matrix(CervicalCancerTest))

TAcc_RF_CervicalC<-max(RFmodel_CervicalC$results$Accuracy)

CM_RF_CervicalC<-table(as.factor(CervicalCancerTest$Biopsy),CervicalCPredicted_RF)
Acc_RF_CervicalC <- sum(diag(CM_RF_CervicalC)) / sum(CM_RF_CervicalC)

```

## Decision Trees (DT)

```{r message=FALSE, warning=FALSE}
###################DT################################


fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid") 

minbucket=0
DTtuneGrid<-expand.grid(cp = c(0.001,0.01, 0.1))

Results_DT_CervicalC<-data.frame()

for (minbucket in c(5,10,15)){
  
  DTmodel_CervicalC <- train(factor(Biopsy) ~ ., 
                           data=as.matrix(CervicalCancerTrain), 
                           method="rpart", 
                           metric="Accuracy",
                           minbucket =minbucket, 
                           trControl=fitControl,
                           tuneGrid = DTtuneGrid)
  
  AccCpMinbucket<-DTmodel_CervicalC$results[,c("Accuracy","cp")]
  AccCpMinbucket$minbucket<-minbucket
  Results_DT_CervicalC<-rbind(AccCpMinbucket,Results_DT_CervicalC)
}

res<-Results_DT_CervicalC%>%arrange(desc(Accuracy))%>%head(1)
res

DTtuneGrid <-  expand.grid(cp =res$cp)

DTmodel_CervicalC_best <- train(factor(Biopsy) ~ ., 
                           data=as.matrix(CervicalCancerTrain), 
                           method="rpart", 
                           metric="Accuracy",
                           minbucket =res$minbucket, 
                           trControl=fitControl,
                           tuneGrid = DTtuneGrid)



CervicalCPredicted_DT<-predict(DTmodel_CervicalC_best,as.matrix(CervicalCancerTest))

TAcc_DT_CervicalC<-min(DTmodel_CervicalC_best$results$Accuracy)

CM_DT_CervicalC<-table(as.factor(CervicalCancerTest$Biopsy),CervicalCPredicted_DT)

Acc_DT_CercivalC<- sum(diag(CM_DT_CervicalC)) / sum(CM_DT_CervicalC)


```

## Stochastic Gradient Boosting (SGB)

```{r message=FALSE, warning=FALSE}
#####################GBM#############################

fitcontrol =trainControl(method="cv", number=5, search="grid")

GBMtuneGrid=expand.grid(interaction.depth = c(2, 4, 6), 
                        shrinkage = c(0.05,0.1, 0.15),
                        n.trees = c(100, 150, 250),
                        n.minobsinnode = 15)

GBMmodel_CervicalC <- train(as.factor(Biopsy) ~ ., 
                            data=as.matrix(CervicalCancerTrain), 
                            method="gbm", 
                            metric="Accuracy", 
                            trControl=fitcontrol,
                            tuneGrid = GBMtuneGrid)


CervicalCPredicted_GBM<-predict(GBMmodel_CervicalC,CervicalCancerTest)

TAcc_GBM_CervicalC<-max(GBMmodel_CervicalC$results$Accuracy)

CM_GBM_CervicalC<-table(as.factor(CervicalCancerTest$Biopsy),CervicalCPredicted_GBM)
Acc_GBM_CervicalC <- sum(diag(CM_GBM_CervicalC)) / sum(CM_GBM_CervicalC)
GBMmodel_CervicalC$bestTune
#print(as.matrix(GBMmodel_HDigits$results))
plot(GBMmodel_CervicalC)



```


## Comparison of the Results

**PRA**

*Cross Validation Error*

```{r message=FALSE, warning=FALSE}
CVError_PRA_CervicalC

```

*Accuracy-Train*

```{r message=FALSE, warning=FALSE}
Acc_PRA_CervicalC_Train
```

*Accuracy-Test*

```{r message=FALSE, warning=FALSE}
Acc_PRA_CervicalC
```

*Lambda*

```{r}
PRAmodel_CervicalC$lambda.min
```


**RF**

*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_RF_CervicalC
```

*Accuracy-Test* 

```{r message=FALSE, warning=FALSE}
Acc_RF_CervicalC

```

*m*

```{r}
RFmodel_CervicalC$bestTune
```


**DT**

*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_DT_CervicalC

```

*Accuracy-Test* 


```{r message=FALSE, warning=FALSE}
Acc_DT_CercivalC
```

*Cp and Minbucket*

```{r}
res$cp
res$minbucket
```


**SGB**
 
*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_GBM_CervicalC
```

*Accuracy-Test* 

```{r message=FALSE, warning=FALSE}
Acc_GBM_CervicalC
```

*Depth, Learning Rate and Number of Trees*

```{r}
GBMmodel_CervicalC$bestTune
```

