---
title: "IE582 Fall"
author: "Fatma Nur Dumlupınar"
date: "29 01 2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    code_folding: hide
    theme: journal
  pdf_document:
    toc: yes
    toc_depth: '3'
subtitle: Homework 4
---

```{r message=FALSE, warning=FALSE}
library(data.table)
library(farff)
library(glmnet)
library(caTools)
library(caret)
library(e1071)
library(MLmetrics)
library(dplyr)
```

# Optical Recognition of Handwritten Digits Data Set

```{r message=FALSE, warning=FALSE}
HandwrittenDigits_test<-fread("data/Optical Recognition of Handwritten Digits Data Set/optdigits_tes.txt")
HandwrittenDigits_train<-fread("data/Optical Recognition of Handwritten Digits Data Set/optdigits_tra.txt")
str(HandwrittenDigits_test)  ## multiclass,65
HandwrittenDigits_test<-HandwrittenDigits_test[,-c(1,40)]
HandwrittenDigits_train<-HandwrittenDigits_train[,-c(1,40)]

```

## Penalized Regression Approaches (RPA)

```{r message=FALSE, warning=FALSE}
PRAmodel_hdigits<- cv.glmnet(as.matrix(HandwrittenDigits_train[,1:62]),
                             as.factor(HandwrittenDigits_train$V65), 
                             alpha = 1,nfolds=5,
                             type.measure='class',family="multinomial")
PRAmodel_hdigits

plot(PRAmodel_hdigits)

HDigitsPredicted_RPA_Train<-predict(PRAmodel_hdigits,as.matrix(HandwrittenDigits_train[,1:62]),
                              type = "class",s=PRAmodel_hdigits$lambda.min)

HDigitsPredicted_RPA<-predict(PRAmodel_hdigits,as.matrix(HandwrittenDigits_test[,1:62]),
                              type = "class",s=PRAmodel_hdigits$lambda.min)

CVError_PRA_HDigits<-min(PRAmodel_hdigits$cvm)

CM_PRA_HDigits<-table(as.factor(HandwrittenDigits_test$V65),HDigitsPredicted_RPA)

Acc_PRA_HDigits <- sum(diag(CM_PRA_HDigits)) / sum(CM_PRA_HDigits)

CM_PRA_HDigits_Train<-table(as.factor(HandwrittenDigits_train$V65),HDigitsPredicted_RPA_Train)

Acc_PRA_HDigits_Train <- sum(diag(CM_PRA_HDigits_Train)) / sum(CM_PRA_HDigits_Train)
```

## Random Forests (RF)


```{r message=FALSE, warning=FALSE}
###Random Forest###################################################


fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid")    

RFtuneGrid=expand.grid(mtry = c(5,15,25))
# the minimal number of observations per tree leaf=5, number of tree J=500,
RFmodel_HDigits <- train(factor(V65) ~ ., 
                         data=as.matrix(HandwrittenDigits_train), 
                         method="rf",
                         metric="Accuracy",
                         ntree =500,nodesize=5,
                         trControl=fitControl,
                         tuneGrid = RFtuneGrid)


HDigitsPredicted_RF<-predict(RFmodel_HDigits,as.matrix(HandwrittenDigits_test))

TAcc_RF_HDigits<-max(RFmodel_HDigits$results$Accuracy)

CM_RF_HDigits<-table(as.factor(HandwrittenDigits_test$V65),HDigitsPredicted_RF)
Acc_RF_HDigits <- sum(diag(CM_RF_HDigits)) / sum(CM_RF_HDigits)


```

## Decision Trees (DT)


```{r message=FALSE, warning=FALSE}
### Decision Tree####################################################

#H Digits

fitControl=trainControl(method = "cv",
                        number = 5,
                        search="grid") 

minbucket=0
DTtuneGrid<-expand.grid(cp = c(0.001,0.01, 0.1))

Results_DT_HDigits<-data.frame()
  
for (minbucket in c(5,10,15)){
  
  DTmodel_HDigits <- train(factor(V65) ~ ., 
                           data=as.matrix(HandwrittenDigits_train), 
                           method="rpart", 
                           metric="Accuracy",
                           minbucket =minbucket, 
                           trControl=fitControl,
                           tuneGrid = DTtuneGrid)
  
  AccCpMinbucket<-DTmodel_HDigits$results[,c("Accuracy","cp")]
  AccCpMinbucket$minbucket<-minbucket
  Results_DT_HDigits<-rbind(AccCpMinbucket,Results_DT_HDigits)
}

res<-Results_DT_HDigits%>%arrange(desc(Accuracy))%>%head(1)
res

DTtuneGrid <-  expand.grid(cp =res$cp)

DTmodel_HDigits_best <- train(factor(V65) ~ ., 
                           data=as.matrix(HandwrittenDigits_train), 
                           method="rpart", 
                           metric="Accuracy",
                           minbucket =res$minbucket, 
                           trControl=fitControl,
                           tuneGrid = DTtuneGrid)



HDigitsPredicted_DT<-predict(DTmodel_HDigits_best,as.matrix(HandwrittenDigits_test))

TAcc_DT_HDigits<-min(DTmodel_HDigits_best$results$Accuracy)

CM_DT_HDigits<-table(as.factor(HandwrittenDigits_test$V65),HDigitsPredicted_DT)

Acc_DT_HDigits <- sum(diag(CM_DT_HDigits)) / sum(CM_DT_HDigits)

```

## Stochastic Gradient Boosting (SGB)


```{r message=FALSE, warning=FALSE}
### Gradient Boosting###################################################

fitcontrol =trainControl(method="cv", number=2, search="grid")

GBMtuneGrid=expand.grid(interaction.depth = c(2, 4, 6), 
                        shrinkage = c(0.05,0.1, 0.15),
                        n.trees = c(100, 150, 250),
                        n.minobsinnode = 15)

GBMmodel_HDigits <- train(factor(V65) ~ ., 
                          data=as.matrix(HandwrittenDigits_train), 
                          method="gbm", 
                          metric="Accuracy", 
                          trControl=fitcontrol,
                          tuneGrid = GBMtuneGrid)


HDigitsPredicted_GBM<-predict(GBMmodel_HDigits,HandwrittenDigits_test)

TAcc_GBM_HDigits<-max(GBMmodel_HDigits$results$Accuracy)

CM_GBM_HDigits<-table(as.factor(HandwrittenDigits_test$V65),HDigitsPredicted_GBM)
Acc_GBM_HDigits <- sum(diag(CM_RF_HDigits)) / sum(CM_RF_HDigits)
GBMmodel_HDigits$bestTune
#print(as.matrix(GBMmodel_HDigits$results))
plot(GBMmodel_HDigits)



```

## Comparison of the Results

**PRA**

*Cross Validation Error*

```{r message=FALSE, warning=FALSE}
CVError_PRA_HDigits

```

*Accuracy-Train*

```{r message=FALSE, warning=FALSE}
Acc_PRA_HDigits_Train
```

*Accuracy-Test*

```{r message=FALSE, warning=FALSE}
Acc_PRA_HDigits
```

*Lambda*

````{r message=FALSE, warning=FALSE}
PRAmodel_hdigits$lambda.min
```

**RF**

*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_RF_HDigits
```

*Accuracy-Test* 

```{r message=FALSE, warning=FALSE}
Acc_RF_HDigits

```

*m*

````{r message=FALSE, warning=FALSE}
RFmodel_HDigits$bestTune
```

**DT**

*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_DT_HDigits

```

*Accuracy-Test* 


```{r message=FALSE, warning=FALSE}
Acc_DT_HDigits
```

*Cp and Minbucket*

````{r message=FALSE, warning=FALSE}
res$cp
res$minbucket
```

**SGB**
 
*Accuracy-Train* 

```{r message=FALSE, warning=FALSE}
TAcc_GBM_HDigits
```

*Accuracy-Test* 

```{r message=FALSE, warning=FALSE}
Acc_GBM_HDigits
```

*Depth, Learning Rate and Number of Trees*

````{r message=FALSE, warning=FALSE}
GBMmodel_HDigits$bestTune
```

